---
title: "HEALTH INSURANCE PREDICTION"
output:
  html_document:
    df_print: paged
---

```{r}
#loading the dataset

library(readr)
insurance<-
  read_csv("C:/Users/mohanraj/Desktop/insurance.csv")
```

```{r}
#summary 

head(insurance)
summary(insurance)
```

```{r}
skimr::skim(insurance)
```

```{r}
#Outlier detection and treatment

library(dlookr)
diagnose_outlier(insurance)
plot_outlier(insurance)
boxplot(insurance$bmi)
outlier_bmi<-boxplot.stats(insurance$bmi)
outlier_bmi$out
```
outliers are detected in variables bmi and charges. Since our response variable is 
charges, we treat outliers only in bmi. 

```{r}
#Treating outlier in bmi by capping

library(tidyverse)
lower_cap<-quantile(insurance$bmi,0.10)
upper_cap<-quantile(insurance$bmi,0.90)
insurance %>% 
  mutate(bmi=if_else(bmi<lower_cap,lower_cap,(if_else(bmi>upper_cap,upper_cap,bmi))))->insurance
```


```{r}
#missing values detection and treatment

library(naniar)
miss_var_summary(insurance)
gg_miss_var(insurance)
```
There is no  missing values in any feature.

```{r}
ggplot(insurance)+geom_bar(aes(x=smoker,fill=region),color='black',position = 'dodge')+facet_wrap(~region,scales = "free_y",nrow = 4)+coord_flip()+theme_minimal()
```
Number of smokers and non smokers almost follows the same pattern across the region

```{r}
boxplot(charges~sex,data=insurance,main='charges across Sex',col=c("green","blue"),horizontal = T)
```
Costs billed by health insurance for 50% of women and men fall under same amount of charge & among the rest the cost billed for men is higher than women.

```{r}
boxplot(charges~smoker,data=insurance,main='charges across Smoker',
        col=c("violet","brown"))
ggplot(insurance,aes(x=age,y=charges))+geom_col(aes(fill=age))+facet_wrap(~smoker)+geom_tile()
  
```
Health insurance billed is compartively much lower for non-smokers than smokers. 
For non-smokers the medical costs rises steadily with increase in age from age 20 to 55.

```{r}
ggplot(insurance,aes(x=bmi,y=charges,col=smoker))+geom_point(aes(shape=smoker),size=2)+geom_smooth(se=F)+theme_grey()
```
The medical cost insurance increases for smoker With increase in bmi .

```{r}
#train-test split

library(tidymodels)
library(bestNormalize)
initial_split(insurance,strata = smoker)->split_data
training(split_data)->train_data
testing(split_data)->test_data
```

```{r}
#model-specification

lin_mod<-linear_reg(mode = "regression")
knn_mod<-nearest_neighbor(mode = 'regression',engine = "kknn",neighbors = tune())
```

```{r}
#recipe for each model

lin_res<-train_data %>% recipe(charges~.) %>% 
  step_dummy(all_nominal_predictors()) %>%          #qualitative variable encoding
  step_zv(all_predictors()) %>%                     #columns with single unique value removed
  step_corr(all_numeric_predictors())               #removing variable with large absolute correlation with other variables

knn_res<-train_data %>% recipe(charges~.) %>% 
  step_dummy(all_nominal_predictors()) %>%         
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors()) %>%              #centred and scaled
  step_best_normalize(all_numeric_predictors())     #transforming
```

```{r}
#cross-validatation data

train_resamples<-vfold_cv(train_data,strata = smoker)
```

```{r}
wflow_train<-workflow_set(preproc = list(lin_res,knn_res),models = list(lin_mod,knn_mod),cross = FALSE)
wmap_train<-wflow_train %>% workflow_map("tune_grid",resamples=train_resamples,grid=25,control=control_grid(save_workflow = T,save_pred = T))
wmap_train 
```

```{r}
#finding the best model among the models

best_workflow_among_set<-wmap_train %>% autoplot(select_best=TRUE,rank_metric='rmse',metric='rmse')
best_workflow_among_set
```
knn model best fits compared to linear model

```{r}
#best parameter selection

best_result<-wmap_train %>% extract_workflow_set_result("recipe_2_nearest_neighbor") %>% select_best('rmse')
best_result
```

```{r}
#fitting

fitted_model<-wmap_train %>% extract_workflow("recipe_2_nearest_neighbor") %>% finalize_workflow(best_result) %>% fit(train_data)
fitted_model
```

```{r}
#predicting 

predicted_test<-fitted_model %>% predict(test_data) %>% bind_cols(test_data %>% select(charges))
predicted_test
```

```{r}
#evaulating the prediction

predicted_test %>% rmse(truth=charges,.pred)
```

