---
title: "HEART FAILURE PREDICTION"
output:
  html_document:
    df_print: paged
---
#dataset variable description

age - Age of patients\n
anaemia - Decrease of red blood cells or hemoglobin (boolean)\n
creatinine_phosphokinase - Level of the CPK enzyme in the blood (mcg/L)\n
diabetes - f the patient has diabetes (boolean)\n
ejection_fraction - Percentage of blood leaving the heart at each contraction\n (percentage)
high_blood_pressure - If the patient has hypertension (boolean)\n
platelets - Platelets in the blood (kiloplatelets/mL)\n
serum_creatinine - Level of serum creatinine in the blood (mg/dL)\n
serum_sodium - Level of serum sodium in the blood (mEq/L)\n
sex - Woman or man (binary)\n
smoking - If the patient smokes or not (boolean)\n
time - Follow-up period (days)\n
DEATH_EVENT - If the patient deceased during the follow-up period (boolean)




```{r}
#loading dataset

library(readr)
heart<-
read_csv(file = "C:/Users/mohanraj/Desktop/heart_failure_clinical_records_dataset.csv")
```

```{r}

head(heart)
tail(heart)
```

```{r}
#checking for incorrect data types and unusal values

summary(heart)
skimr::skim(heart)

```
There is no missing value in any variable. There seems to unusal value in creatinine_phlosphokinase.

```{r}
#outlier detection and treatment

library(dlookr)
library(tidyverse)
select(heart,find_outliers(heart)) %>% plot_outlier()
select(heart,find_outliers(heart)) %>% diagnose_outlier()

```

```{r}
#capping

heart<-
heart %>% 
  mutate(
    creatinine_phosphokinase=imputate_outlier(heart,xvar = creatinine_phosphokinase,method="capping",no_attrs = T),
    ejection_fraction=imputate_outlier(heart,xvar =ejection_fraction,method="capping",no_attrs = T),
    platelets=imputate_outlier(heart,xvar =platelets,method="capping",no_attrs = T),
    serum_creatinine=imputate_outlier(heart,xvar = serum_creatinine,method="capping",no_attrs = T),
    serum_sodium=imputate_outlier(heart,xvar = serum_sodium,method="capping",no_attrs = T))
```

```{r}
#correlation plot
correlation=cor(heart)
corrplot::corrplot(correlation,method = "color")
```
correlation plot indicates absence of correlation among the explanatory variables.\n
Predictor variable death_event is negative correlated with time (i.e) possibility of dying increases with decrease in time\n
From the plot it seems sex ,smoking and diabetes shows no correlation with death event\n
lets further investigate

```{r}
#frequency of death_event across sex, smoking and diabetes

sex_deathevent<-heart %>% group_by(sex) %>% count(DEATH_EVENT)
smoking_deathevent<-heart %>% group_by(smoking) %>% count(DEATH_EVENT)
diabetes_deathevent<-heart %>% group_by(diabetes) %>% count(DEATH_EVENT)
```

```{r}
#independence test 

chisq.test(sex_deathevent)
chisq.test(smoking_deathevent)
chisq.test(diabetes_deathevent)
```
Sex, diabetes and smoking are not statistically associated with death event

```{r}
#finding correlation significance of sex and death event with death event

cor.test(heart$sex,heart$DEATH_EVENT)
cor.test(heart$smoking,heart$DEATH_EVENT)
cor.test(heart$diabetes,heart$DEATH_EVENT)
```
The p>0.05 for both tests. Therefore, it is confirmed that \n sex ,smoke and diabetes are not correlated with death event.

```{r}
#train-test split and validation 

library(tidymodels)
heart<-heart %>% mutate(DEATH_EVENT=as.factor(if_else(DEATH_EVENT==0,"0","1")))
initial_split(heart,strata=DEATH_EVENT)->split
training(split)->train
testing(split)->test

```

```{r}
#model-specification

logistic_mod<-logistic_reg("classification")
```

```{r}
#preprocessing and formula

logistic_rec<-heart %>% recipe(DEATH_EVENT~.) %>% 
  step_rm(sex,smoking,diabetes) %>% 
  step_dummy(all_nominal_predictors()) %>%   
  step_zv(all_predictors()) %>% 
  step_corr(all_predictors())

```

```{r}
#fitting model

fitting<-workflow() %>% add_recipe(logistic_rec) %>% add_model(logistic_mod) %>% fit(train)
```

```{r}
#predicting

prediction<-fitting %>% predict(test) %>% bind_cols(test %>% select(DEATH_EVENT))
prediction
```

```{r}
#confusion matrix and prediction accuracy

prediction %>% conf_mat(truth=DEATH_EVENT,.pred_class)
prediction %>% accuracy(truth=DEATH_EVENT,.pred_class)

```
Accuracy is 0.8533. Therefore, the model accurately predicted for 85% of survival of patients.\n


```{r}
#roc_curve
fitting %>% predict(test,type="prob") %>% bind_cols(test %>% select(DEATH_EVENT))%>% roc_curve(truth=DEATH_EVENT,.pred_0) %>% autoplot()
```

